---
title: "BDA"
author: "Julian Frattini, Andreas Bauer"
output: html_document
date: "2023-06-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggdag)
```

This document contains the analysis of the effect of augmented testing on the relative duration for GUI testing.

## Causal Assumptions

### Hypotheses

We formulate the following hypotheses based on our prior knowledge:

1. The use of augmented testing has an influence on the relative duration for GUI testing.
2. The size of a test case has an influence on the relative duration for GUI testing.

Additionally, we need to consider that *the repeated use of the augmented testing system has an influence on the relative duration for GUI testing* (learning effect).

### Directed Acyclic Graph

We can visualize these hypotheses in the following graph:

```{r dag} 
dag <- dagify(
  dur ~ at + size + learn,
  exposure = "at",
  outcome = "dur",
  labels = c(dur = "duration_scaled", at = "augmented_testing", size="testcase_size", learn="learning_effect"),
  coords = list(x=c(at=0, size=0, learn=0, dur=2),
                y=c(at=1, size=0, learn=-1, dur=0))
)

ggdag_status(dag, 
             use_labels = "label", 
             text = FALSE) +
  guides(fill = "none", color = "none") + 
  theme_dag()
```


## Data

First, load the data from the table of results.

```{r load-data}
df_raw <- read.csv(file = '../data/results.csv', header = TRUE, sep = ',', fileEncoding="UTF-8-BOM")
```

### Cleanup and transform data

The data has to be transformed into the following format

| Column | Description | Type |
|---|---|---|
| `id` | The identifier of a participant | int |
| `tc` | The identifier of a test case | int |
| `tc_big` | Size of the test case.  `TRUE` for bigger and `FALSE` for smaller test cases. | Boolean |
| `treatment` | `TRUE` for Augmented Testing and `FALSE` for manual GUI testing | Boolean |
| `duration_scaled` | Duration of performing test cases (normalized) | float |

To achieve this, first [pivot](https://tidyr.tidyverse.org/reference/pivot_longer.html) the table from wide to long format.

```{r pivot}
df <- df_raw %>% 
  pivot_longer(
    cols = c(matches("TC._treatment"), matches("TC._seconds$")),
    names_to = c("tc", ".value"), names_pattern = "TC(.)_(.*)"
  ) %>% 
  select("ID", "tc", "treatment", "seconds") %>% 
  mutate(
    treatment = (treatment == "A"),
    tc_big = (tc %in% c(3, 4, 7, 8))
  )

head(df)
```

Next, normalize the duration values: scale the duration to an interval of size 1 centered around the mean value of each test case.

```{r scale}
df <- df %>% 
  group_by(tc) %>% 
  mutate(
    duration_scaled = (seconds-mean(seconds))/(max(seconds)-min(seconds))
  )
```

Finally, add the learning variable: because the order of the test cases was the same for all participants, there is a potential learning effect in using AT. Hence, determine the *number of previous test cases in which augmented testing has already been used* to represent the learning effect.

```{r learning}
# determine the group of participants which obtained the treatment first, i.e., all with an even ID
group.treatment.first <- seq(2, max(df$ID), 2)

# for each of the two groups, determine the "number of previous test cases employing augmenting testing"
df <- df %>% 
  mutate(
    at.learned = ifelse(
      ID %in% group.treatment.first, 
      ifelse(tc %in% c(1,3,5,7), (as.integer(tc)-1)/2, 4),
      ifelse(tc %in% c(4,6,8), (as.integer(tc)-2)/2, 0))
  )
```

## Modeling

Now, model the data using Bayesian data analysis.

```{r bda-lib}
library(brms)
library(patchwork)
library(marginaleffects)
```

### Formula

First, we define our models by (1) representing our DAG - i.e., the assumed causal relationships within our data - in a formula, and (2) selecting an appropriate distribution. In this analysis, we will compare the following models, which are further explained below:

| ID | Distribution | Interaction | Learning | F |
|---|---|---|---|---|
| `m1` | Gaussian | no | no | `f1` |
| `m2` | Gaussian | yes | no | `f2` |
| `m3` | Gaussian | yes | yes | `f3` |
| `m4` | Skew_normal | yes | yes | `f3` |

For each formula, we can obtain a list of prior parameters which need to be defined later on via `brms::get_prior`.

The model `m1` simply assumes an influence of the `treatment` and the test case size (`tc_big`) on the `duration_scaled`.

```{r formula-standard}
f1 <- duration_scaled ~ treatment + tc_big

brms::get_prior(f1, family=gaussian, data=df)
```

The model `m2` additionally assumes an [interaction effect](https://bookdown.org/content/3890/interactions.html) between the two predictors.

```{r formula-interaction}
f2 <- duration_scaled ~ treatment * tc_big

brms::get_prior(f2, family=gaussian, data=df)
```

The model `m3` additionally assumes a learning effect which occurs as an interaction between the treatment and the number of test cases already processed using augmented testing (`at.learned`). The hypothesis is that the more often a subject has used the augmented testing system, the quicker they become (i.e., the lower the `duration_scaled` becomes).

```{r formula-learning}
f3 <- duration_scaled ~ treatment * tc_big + treatment * at.learned

brms::get_prior(f3, family=gaussian, data=df)
```

The model `m4` assumes that the response variable `duration_scaled` does not follow a Gaussian distribution, but rather a [skewed Gaussian distribution](http://paul-buerkner.github.io/brms/reference/SkewNormal.html). We derive this insight from the posterior predictive checks.

```{r formula-skew}
brms::get_prior(f3, familiy=skew_normal, data=df)
```

All of these models represent our DAG but with slightly different ontological assumptions. We will compare the predictive power of these models to determine, which of these models explains the data best.

### Priors

Next, select sensible priors. We need to set priors for the `Intercept`, the slope `b` (or "beta"), and the standard deviation `sigma`.

```{r priors}
priors <- c(
  prior(normal(0, 0.5), class=Intercept),
  prior(normal(0, 0.5), class=b),
  prior(weibull(2, 1), class=sigma)
)

priors2 <- c(
  prior(normal(0, 0.3), class=Intercept),
  prior(normal(0, 0.3), class=b),
  prior(weibull(2, 1), class=sigma)
)
```

To assess the feasibility of the selected priors, sample from only the priors (i.e., avoid the Bayesian learning process for now) by setting `sample_prior="only"`.

```{r m1-prior-sampling}
m1.prior <- brm(
  data = df, # provide the data
  family = gaussian, # determine the distribution type of the response variable (here: Gaussian)
  formula = f1, # provide the formula
  prior = priors, # provide the priors
  iter = 4000, warmup = 1000, chains = 4, cores = 4, seed = 4, # brms settings
  sample_prior="only", # do not train the parameters, sample only from the priors, 
  file = "fits/m1.prior" # save the fit in a file
)
```

```{r m2-prior-sampling}
m2.prior <- brm(
  data = df, family = gaussian,
  formula = f2, prior = priors,
  iter = 4000, warmup = 1000, chains = 4, cores = 4,
  seed = 4, sample_prior="only", file = "fits/m2.prior"
)
```

```{r m3-prior-sampling}
m3.prior <- brm(
  data = df, family = gaussian,
  formula = f3, prior = priors2,
  iter = 4000, warmup = 1000, chains = 4, cores = 4,
  seed = 4, sample_prior="only", file = "fits/m3.prior"
)
```

```{r m4-prior-sampling}
m4.prior <- brm(
  data = df, family = skew_normal,
  formula = f3, prior = priors2,
  iter = 4000, warmup = 1000, chains = 4, cores = 4,
  seed = 4, sample_prior="only", file = "fits/m4.prior"
)
```

With the competing models defined and sampled from the priors, we can perform a [graphical prior predictive check](https://mc-stan.org/bayesplot/reference/PPC-overview.html) by visualizing the sampled values against the actually observed values.

```{r prior-predictive-check}
m1.priorpc <- brms::pp_check(m1.prior, type="dens_overlay_grouped", ndraws=100, group="treatment") + ggplot2::ggtitle("M1")
m2.priorpc <- brms::pp_check(m2.prior, type="dens_overlay_grouped", ndraws=100, group="treatment") + ggplot2::ggtitle("M2")
m3.priorpc <- brms::pp_check(m3.prior, type="dens_overlay_grouped", ndraws=100, group="treatment") + ggplot2::ggtitle("M3")
m4.priorpc <- brms::pp_check(m4.prior, type="dens_overlay_grouped", ndraws=100, group="treatment") + ggplot2::ggtitle("M4")

m1.priorpc / m2.priorpc / m3.priorpc / m4.priorpc
```

We expect that the sampled values ($y_{rep}$) are close to the actually observed values ($y$), which confirms that the actual data is realistic in the eyes of the model given its prior believes.

### Training

With sufficiently sensitive priors, we can train the model on the actual data. For this, we remove the `sample_prior="only"` argument. Instead, we add the `file=fits/m1` argument such that the trained model will be saved in the "fits" folder within the "src" folder (make sure that this folder already exists).

```{r m1-training}
m1 <- brm(
  data = df, family = gaussian,
  formula = f1, prior = priors,
  iter = 4000, warmup = 1000, chains = 4, cores = 4,
  seed = 4, file = "fits/m1"
)
```

```{r m2-training}
m2 <- brm(
  data = df, family = gaussian,
  formula = f2, prior = priors,
  iter = 4000, warmup = 1000, chains = 4, cores = 4,
  seed = 4, file = "fits/m2"
)
```

```{r m3-training}
m3 <- brm(
  data = df, family = gaussian,
  formula = f3, prior = priors2,
  iter = 4000, warmup = 1000, chains = 4, cores = 4,
  seed = 4, file = "fits/m3"
)
```

```{r m4-training}
m4 <- brm(
  data = df, family = skew_normal,
  formula = f3, prior = priors2,
  iter = 4000, warmup = 1000, chains = 4, cores = 4,
  seed = 4, file = "fits/m4"
)
```

To confirm that the training was successful, we can perform a [graphical posterior predictive check](https://mc-stan.org/bayesplot/reference/PPC-overview.html) similar to the prior predictive check, but this time sampling from the actually trained models

```{r posterior-check}
m1.postpc <- brms::pp_check(m1, type="dens_overlay_grouped", ndraws=100, group="treatment")
m2.postpc <- brms::pp_check(m2, type="dens_overlay_grouped", ndraws=100, group="treatment")
m3.postpc <- brms::pp_check(m3, type="dens_overlay_grouped", ndraws=100, group="treatment")
m4.postpc <- brms::pp_check(m4, type="dens_overlay_grouped", ndraws=100, group="treatment")

m1.postpc / m2.postpc / m3.postpc / m4.postpc
```

The visualizations confirm that the parameters of the trained model imply distributions very close to the actually observed values, but only model `m4` accommodates the skeweness that the values for `treatment = TRUE` imply.

### Model Comparison

To determine, which of the two formulas (and, hence, which of the two models) fit the data better, we can perform a [leave-one-out comparison](http://paul-buerkner.github.io/brms/reference/loo_compare.brmsfit.html). 

```{r loo-compare}
m1 <- add_criterion(m1, criterion = "loo")
m2 <- add_criterion(m2, criterion = "loo")
m3 <- add_criterion(m3, criterion = "loo")
m4 <- add_criterion(m4, criterion = "loo")

loo_compare(m1, m2, m3, m4)
```

The model `m4` clearly outperforms all other models, which is visible through its $elpd\_diff$ value of 0.0, while all other models perform significantly worse. Model `m4` hence has the best predictive power and will be used to move forward.

## Evaluation

We can evaluate the best-fitting model to infer insights about the impact of the modeled predictors.

### Parameters

First, we can take a look at the parameters, which are the priors updated based on the observed data.

```{r parameters}
summary(m4)
```

We see the following effects:

1. A mostly negative distribution of `treatmentTRUE` ("mostly negative" meaning that the most of the 95% confidence interval is negative), which means that employing augmented testing is predominantly associated with a `scaled_duration` value below 0, i.e., below the average for that test case. 
2. A strongly positive distribution of `tc_bigTRUE`, meaning that larger test cases tend to have a longer-than-average time for testing.
3. A strongly negative distribution of `treatmentTRUE:tc_bigTRUE`, meaning that when a test case is large, the usage of augmented testing reduces the testing duration even stronger.
4. A negligibly negative distribution of `treatmentTRUE:at.learned`, meaning that the the learning effect coinciding with the treatment very slightly reduces the scaled duration.

These are the interpretations of the raw parameters which do not account for *all of the uncertainty* which the model has picked up, but isolate the effect of the considered predictors.

### Marginal and Conditional effects

A more sophisticated evaluation uses plots of marginal and conditional effects.

```{r effect-plots}
eff <- conditional_effects(m4)
```

#### Marginal effects

[Marginal effects](https://www.andrewheiss.com/blog/2022/05/20/marginalia/) represent the isolated effect of a predictor while fixing all other predictors at an average (mean or mode) level.

```{r marginal-plots}
marginal.effect.treatment <- plot(eff, plot = FALSE)[[1]]
marginal.effect.size <- plot(eff, plot = FALSE)[[2]]
marginal.effect.learning <- plot(eff, plot = FALSE)[[3]]

marginal.effect.treatment | marginal.effect.size | marginal.effect.learning
```

In our case, it visualizes both the effects but also the uncertainty of the impact of each individual predictor.

#### Conditional effects

Conditional effects visualize the interaction between two predictor variables. In the currently evaluated model, we have two interaction effects to look into:

1. The interaction between the *treatment* and the *test case size* (`treatment * tc_big`)
2. The interaction between the *treatment* and the *learning effect* (`treatment * at.learned`)

```{r conditional-plots}
conditional.effect.treatment <- plot(eff, plot = FALSE)[[4]] + theme(legend.position="bottom")
conditional.effect.learning <- plot(eff, plot = FALSE)[[5]] + theme(legend.position="bottom")

conditional.effect.treatment | conditional.effect.learning
```

The visualizations show the following:

1. Larger test cases benefit much stronger from the use of augmented testing than smaller test cases: when comparing the distributions for `treatment=FALSE` vs `treatment=TRUE`, the distance between the two means is much bigger for `tc_big=TRUE` than `tc_big=FALSE`.
2. The repeated use of augmented testing is slightly beneficial in terms of efficiency, as the effect of `at.learned` is more negative (i.e., reduced `duration_scaled` more) when `treatment=TRUE` than when `treatment=FALSE`.

The second interaction effect is fairly trivial: getting used to the augmented testing system (Scout) has a stronger effect on effectiveness when using augmented testing than when not using it. But even if this insight is trivial, it shows us that the expected learning effect is present in the data and modeling it makes the other predictors more precise by explaining at least this fraction of the effect on the response variable by the learning effect.

### Sampling from the Posterior

Finally, to account for all of the uncertainty of the model, we can sample from the posterior. For this, we perform the following steps:

1. Define a new `datagrid` in which we fix all predictors. In our case, we once set the treatment to `TRUE` and once to `FALSE`. In both cases, we want an even distribution of the other predictor, `tc_big` (hence, we set it to `c(TRUE, FALSE)`).
2. Sample from the model given both new datagrids.
3. Compare the two samples by subtracting the resulting, sampled `duration_scaled` values from each other (`diff`), only keep the sign (i.e., + or -) of the difference (`sign`), and count, how often the sampled `duration_scaled` value from `treatment=TRUE`-sample was greater than from the `treatment=FALSE`-sample (`table`).

Finally, we output those counts divided by the total amount, i.e., the percentage.

```{r posterior-sampling}
posterior.draws.baseline <- posterior_predict(
  m4, newdata = datagrid(
    model = m4,
    treatment = FALSE,
    tc_big = c(TRUE, FALSE)
    ))

posterior.draws.treatment <- posterior_predict(
  m4, newdata = datagrid(
    model = m4,
    treatment = TRUE,
    tc_big = c(TRUE, FALSE)
    ))

diff <- posterior.draws.treatment - posterior.draws.baseline
tab <- table(sign(diff))
  
  
(tab/sum(tab))
```

This means, that - accounting for all uncertainty that the model picked up and under equal distribution of test case size - the use of augmented testing on average results in shorter test duration in about 70% of all cases.
